\documentclass{article}
\usepackage{amsmath,amsfonts,amsthm,fancyhdr,parskip,amssymb,graphicx}
\usepackage[all]{xy}
\usepackage[margin = 1.5in]{geometry}
\pagestyle{fancy}
\lhead{Ben Carriel (bkc39)}
\chead{Bryan Cuccioli (blc72)}
\rhead{Andy Levine (awl58)}
\setlength{\parindent}{1cm}

\DeclareMathOperator{\Z}{\mathbb{Z}}
\DeclareMathOperator{\Q}{\mathbb{Q}}
\DeclareMathOperator{\R}{\mathbb{R}}

\DeclareMathOperator{\oh}{\mathcal{O}}
\DeclareMathOperator{\ta}{\xrightarrow{\ \ \ }}
\DeclareMathOperator{\opt}{\texttt{OPT}}

\newcommand{\problem}[1]{\noindent {\bf #1}}
\newcommand{\problempart}[1]{\noindent{\textbf{(#1)}}}

\newtheorem*{thm}{Theorem}
\newtheorem*{lem}{Lemma}
\newtheorem*{claim}{Claim}
\newtheorem*{defn}{Definition}
\newtheorem*{prop}{Proposition}

\begin{document}

\problem{Problem 1.}

\problempart{a} This is straightforward. The only step in the algorithm that ever changes the capacities of the edges in the residual graph is an augmentation step. But we assumed that all of the edges had $0,1$ capacity and we only subtract values so it is clear that the remaining edges have residual capacity $-1,0,1$. But we couldn't have subtracted $1$ from $0$, because if $e_1$ is a capacity 0 edge, then it will not be in any $s-t$ augmenting path, because no flow can pass through it. Hence, we can only subtract one from edges that had capacity 1 and so we still have $0-1$ edge capacities. 

\problempart{b} Trace some augmenting path (via successive advance steps). The flow through any augmenting path must be 1, and hence each edge in such a path is saturated, and also a bottleneck. This means that every retreat step deletes an edge. The total number of possible deletes is bounded above by the number of edges in the graph. And so we have that the total amount of time to compute a blocking set of flows in $O(m)$.

\problempart{c} Suppose that the value of the max-flow in the graph, $f^*$, is less than $m^{1/2}$ then the result is clear because we have that 
\[
\ell \leq \frac{m}{\ell} \leq m^{1/2}
\]
where $\ell$ is the length of the longest path. Now we consider the case where $f^* > m^{1/2}$. We want to look at the phase where we have equality, that is when $f = f^* - m^{1/2}$. At this stage we have that the flow is $f < f^* - m^{1/2}$ when the level graph for this phase is constructed. Observe that at this point the level graph is the same as the initial level graph in the residual flow network $G_f$ and so we have that the maximum flow in this residual graph is
\[
f^* - f > f^* - (f^* - m^{1/2}) = m^{1/2}
\]
Now we can say the same as the first case
\[
\ell < \frac{m}{f^*} \leq m^{1/2}
\]
At this point we can have gone through at most $m^{1/2} - 1$ and then the number of remaining phases is at most $m^{1/2}$ and so the total number of phases is at most $2m^{1/2} = O(m^{1/2})$.

\problempart{d} Suppose that the max flow in the graph is $f^* \leq n^{2/3}$, then the bound is obvious and has the same argument as in Part(c). If not, we follow the same proof outline as before, at the phase where the flow reaches the value $f^* - n^{2/3}$ we let $f$ be the flow at the point that we construct the level graph for this phase. As before, this level graph is identical to the level graph at the start of the next phase. The one key difference is that now we may have parallel edges between pairs of vertices. This leads us to the following\\

\begin{lem}
Let $f$ be a flow in $G$ and let $f^*$ be the maximum flow in $G$. Then the maximum flow in the next phase of Dinic's algorithm on $G$ is $f^* - f$. 
\end{lem}
\begin{proof}
Let $G'$ be the graph in the next phase of Dinic's algorithm (i.e. what remains after an Augment step). Then we must have that for any cut $k = (A,B)$ with $s\in A, t\in B$ that
\[
\sum_{e\in k'} c'(e) = \left(\sum_{e_1 \in k} c(e_1) - f(e_1)\right) + \sum_{e_2 \in k_{AB}} f(e_2) 
\]
Where $k'$ is the cut after passing the augment step, $c'(e)$ is the new capacity after an augment step (residual capacity) and$k_{AB}$ are the edges in the cut containing vertices in both $A$ and $B$ (edges across the cut). But note that
\[
f = \sum_{e \in k} f(e) - \sum_{e' \in k_{AB}} f(e')
\] 
And so
\[
\sum_{e \in k'} c'(e) = \sum_{e' \in k} c(a) - f
\]
So the minimum cuts are preserved after passing through a phase of the algorithm. We then apply the max-flow min-cut theorem to see that the value of the minimum cut is also $f^*$ in $G'$. So the value of a min-cut in $G'$ is $f^* - f$ and so we have that the maximum flow must also be $f^* - f$ by the min-cut theorem.
\end{proof}

We can apply the lemma to see that the max-flow in the beginning of the next phase is at least $n^{2/3}$. Then we move to the following \\
\begin{lem}
Let $G$ be a flow graph with max flow $f^*$. Then the minimum length $s-t$ when the flow is identically zero is $2\sqrt{2}n/\sqrt{f^*}$
\end{lem}
\begin{proof}
Consider the set $V_d$ of vertices of distance $d$ away from $s$ in $G$. If there are no parallel edges in $G$ then 
\[
f^* \leq |V_d|\cdot|V_{d+1}| 
\]
In the other case we have that there may be twice as many edges so the right wide must be multiplied by a factor of $2$. Then note that either we have
\[
|V_d| \geq 2^{-1/2}(f^*)^{1/2} \text{ or } |V_{d+1}| \geq 2^{-1/2}(f^*)^{1/2}
\]
If $\ell$ is the length of the minimum path then we have 
\[
\sum_{d=0}^{\ell} |V_d| \leq n
\]
which means that
\[
\left\lfloor \frac{l+1}{1}\right\rfloor 2^{-1/2}(f^*)^{1/2} \leq n
\]
Giving $\ell \leq 2\sqrt{2}n/\sqrt{f^*}$
\end{proof}
We then apply the lemma to see that 
\[
\ell \leq \frac{2\sqrt{2}n}{\sqrt{n^{2/3}}} = 2\sqrt{2}n^{2/3}
\]
So the number of phases is at this point was $O(n^{2/3})$ and the number of phases to termination of the algorithm is at most $n^{2/3}$ we have that the total number of phases is $O(n^{2/3})$. As a result of this the total runtime of Dinic's Algorithm on $G$ with $0-1$ edge capacities is
\[
O(m\min\{m^{1/2}, n^{2/3}\})
\]


\problem{Problem 2.}

We will construct a flow network $G$ with every edge having capacity 0 or 1. Assign capacity 1 to all pairs of nodes linked by edges in $G$ and 0 to the other pairs. Then each source in $A\cup B$ can only produce a flow of 1; we specify this by creating a super source $s^{\ast}$ with edges of capacity 1 into each source of $A\cup B$.

Now, augment the network by all augmenting paths starting from nodes in $A$ which give a flow $|A|$. This flow is suboptimal since there is possible flow at least $|B|>|A|$. Let $G_f$ be the residual graph induced by this flow. There is an edge-disjoint path to every node in $A$ from $t$ now that we have augmented these paths.

Moreover, there must now exist some augmenting path $p$, since our flow is sub-optimal. But each node in $A$ has only 1 unit of flow to push, so $p$ must originate from a node in $B\setminus A$. Denote this node $s$. Augmenting by $p$ gives an edge disjoint path from $t$ to every node in $A\cup\{s\}$ in $G_f$, and it follows that in $G$ there are edge-disjoint paths to $t$ for every $n\in A\cup\{s\}$.

\problem{Problem 3.} 

We will prove that an admissible schedule exists by construction of an algorithm. The main strategy is a reduction to maximum flow. As inputs to the problem we are given a set of $n$ towns and $m$ days as well as a matrix of values $g(i,j)$ which measures the amount of garbage produced int town $i$ on day $j$. From these values we can construct a flow network as follows:
\begin{enumerate}
\item  Create a vertex $v_{ij}$ to represent town $i$ on day $j$. That is, one vertex for each entry in the matrix.
\item Create a set of vertices $w_j$ for each $j \leq m$
\item Add a source node $s$ and a sink node $t$
\item Add an edge $e_i = (v_{ij}, w_j)$ for each $i$ and $j$. These are meant to force the truck to go to only one town on each day.
\item We add edges $f_j = (w_j, t)$ for each $w_j$. This forces there to be less than to units of garbage on any given day. 
\item Then we add edges to connect each town inter-day that is $e_v = (v_{ij}, v_{ij+1})$. 
\item Add edges out of the source according to the process described below
\end{enumerate}

All edges in the graph up to this point have capacity 1. We then add edges out of the source to each $v_{ij}$ according to the following capacity function
\[
c(s,v_{ij}) = \sum_{k=j}^mg(i,k) - \sum_{k = j+1}^m g(i,k)
\]
What this formula quantifies is the amount of garbage that we would save if we sent the truck to town $i$ at day $j$. Then we run the max-flow algorithm (any will do) on $G = (\{v_{ij}, w_j\} , E, s, t)$ and return the schedule that corresponds to the incoming edges of $w_j$. That is, we choose to go to town $i$ on day $j$ iff $f(e_i) > 0$. \\

We now proceed to the runtime analysis of the algorithm. The construction of the graph takes $O(mn) + O(n)$ to add the vertices and $O(mn)$ to construct the edges. We will see that the algorithm will produce a max flow via the Edmonds-Karp algorithm in $O((mn)^3)$ and if we restrict the edge wights to $\{0,1\}$, which can be done by taking the floor of each of the edges coming out of the source (the only possibly non-integer capacity edges) and then run Dinic's algorithm for a runtime of 
\[
O(m\min\{m^{1/2}, n^{2/3}\}) = O((m+n)^{3/2})
\]
The significance of the restriction to integer weights will become significant shortly. 

We now move to the proof of correctness. Let $f$ be a max flow in the constructed graph, $G$. We need to show that this flow gives an admissible schedule. In the non-integer weighted graph, we can see that the output of the algorithm will be a maximal fractional flow. and to prove an admissible schedule exists we must turn it into an integer flow. This is clear because if not then there would be some interval $(d_p, d_q)$ such that no garbage is collected in some town and the garbage accumulates above 2. We can see that 
\begin{align*}
\sum_{k = d_p}^{d_q} c(s, v_{pq}) &= \sum_{k = d_p}^{d_q} \left(\sum_{r=k}^m g(p,r) - \sum_{r = k+1}^m g(i,r)\right) \\ 
\end{align*}1
Cancelling common terms yields
\[
\sum_{r=p}^m g(p,r) - \sum_{r = q+1}^m g(q,r)
\]
Apply the Cauchy-Schwartz inequality to see that
\[
\sum_{r=p}^m g(p,r) - \sum_{r = q+1}^m g(q,r) \geq \sum_{r=p}^m g(p,r) \geq 2
\]
But then we have a set of vertices such that there is at least two flow out of the source ,and only has one edge leading to the sink. So the maximum flow can be bounded by
\[
\sum_{i,j} c(s, v_{ij}) - 1
\]
Which is admissible. Next we remark that we can change these fractional weights to integer ones via the floor function because we have that the amount of trash out of a $w_j$ commits one total fractional truck per day, and then we note that it will choose town $i$ as soon as the total weighted truck is 1. But then we see that this happens precisely when there would be 1 flow in the integer graph. So we are done. 
\end{document}